---
  output: html_document
---
  
##### Jongchul Seon


##### We are going to compare genes which shows significant relationships with young and old which can explain biological aging predicting cognitive function. The original paper is http://www.ncbi.nlm.nih.gov/pubmed/26343147. It is only the first test on the paper. 

##### The profiling was done using the Affymetrix array. We are going to use oligo package.


```{r chunk1,echo=TRUE}

library(GEOquery)
library(oligo)
library(dendextend)
library(biomaRt)
library(limma)

##
getwd()

#getGEOSuppFiles("GSE59880")
##list.files("GSE59880")
##untar("GSE59880/GSE59880_RAW.tar", exdir = "GSE59880/CEL")
##list.files("GSE59880/CEL")

celfiles <- list.files("D:/R/GSE59880/CEL", full = TRUE)
rawData <- read.celfiles(celfiles)


###library(pd.hg.u133.plus.2)

```


##### raw data 
##### We can see some variances in old according to the data. However, I assume that there is no big expression variances among old and will use rma for normalization.

```{r chunk2,echo=TRUE}

edata_raw <- exprs(rawData)

dim(edata_raw)

edata_raw[1:10,]

sum(is.na(edata_raw))

par(mfrow=c(1,2))

boxplot(log(edata_raw[,1:15]), main="old", col=2,range=0 )

boxplot(log(edata_raw[,16:30]), main="young", col=2,range=0 )

par(mfrow=c(1,1))

hist(log(edata_raw[,1]))

mm = log(edata_raw[,1]+1) - log(edata_raw[,2]+1)
aa = log(edata_raw[,1]+1) + log(edata_raw[,2]+1)
plot(aa,mm,col=2)


```



##### normalization

```{r chunk3,echo=TRUE}

normData <- rma(rawData)
normData

dim(normData)

head(exprs(normData))

edata <- exprs(normData)

boxplot(edata[,1:30],main="expression data",col=2)

```


##### phenotype data


```{r chunk4,echo=TRUE}

geoMat <- getGEO("GSE59880",destdir = "D:\\R")

### destdir = "D:\\R" => window OS directory

pD.all <- pData(geoMat[[1]])

head(pD.all)
names(pD.all)

##pD <- pD.all[, c("title", "geo_accession", "characteristics_ch1.1","characteristics_ch1.2","characteristics_ch1.3","characteristics_ch1.4","platform_id"  )]


pD <- pD.all[, c("title", "geo_accession", "characteristics_ch1.1","characteristics_ch1.2")]

names(pD)[c(3,4)] <- c("age", "fitness")
## head(pD)

##pD$age
pD$age <- sub("^age: ","", pD$age)
##pD$age

##pD$fitness 
pD$fitness <- sub("^aerobic fitness: ", "", pD$fitness)
##pD$fitness

changeName <- function(x) {
  
    print("function starts.... object length is")
    print(length(x))
    
    for (i in 1:length(x)) {
    
      x[i] <- substr(x[i], start=1, stop=10) 
      
      print(x[i])
    }
     
  return(x)
   
}

sampleNames(normData) <- changeName(sampleNames(normData))

sampleNames(normData)

### rownames(pD) 

group <- ifelse((pD$age)>50, "old", "young")

pD <- cbind(pD,group)

pData(normData) <- pD
```



```{r chunk5,echo=TRUE}


mod = model.matrix(~ pData(normData)$group)
fit_limma = lmFit(exprs(normData),mod)
ebayes_limma = eBayes(fit_limma)


```


```{r chunk6,echo=TRUE}


limma_pvals = topTable(ebayes_limma,adjust.method="BH",sort.by="none",number=dim(edata)[1])

dim(limma_pvals)

head(limma_pvals[limma_pvals$adj.P.Val < 0.05,])

sum(limma_pvals$adj.P.Val < 0.05)

cgenes <- rownames(limma_pvals[limma_pvals$adj.P.Val < 0.05,])

length(cgenes)

###genes <- limma_pvals$adj.P.Val < 0.05

###DE <- featureNames(fdata_bot)[genes]

###length(DE)

```

####select gene expression data strongly related with aging according to the paper. 
 

```{r chunk7,echo=TRUE}

library(xlsx)
data <- read.xlsx("D:/R/13059_2015_750_MOESM1_ESM.xlsx", 1, startRow=2, endRow=152,colIndex=1)

head(data)

genes <- as.vector(as.matrix(data))

```


```{r chunk8,echo=TRUE}

sum(cgenes %in% genes)

```


```{r chunk9,echo=TRUE}

genes_age <- exprs(normData)[rownames(exprs(normData))  %in% genes,]

head(genes_age)

boxplot(genes_age)

dist1 = dist(t(genes_age))

hclust1 = hclust(dist1)
plot(hclust1)

dend = as.dendrogram(hclust1)
dend = color_labels(hclust1,2,col=1:2)
plot(dend)

heatmap(genes_age)

```


```{r chunk10,echo=TRUE}


limma_pvals_test1 = topTable(ebayes_limma,adjust.method="BH",sort.by="logFC",number=dim(edata)[1])

sum(rownames(limma_pvals_test1)  %in% genes)


limma_pvals_test2 = topTable(ebayes_limma,adjust.method="hochberg",sort.by="logFC",number=dim(edata)[1])

tgenes <- rownames(limma_pvals_test2[limma_pvals_test2$adj.P.Val < 0.05,])

sum(tgenes %in% genes)

###affy_probes <- read.xlsx("D:/R/GPL570-13270.xlsx", 1, startRow=20,endRow=500)


```

```{r chunk11,echo=TRUE}

devtools::session_info()

``````